\documentclass[12pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[
    left=2.5cm,
    right=2.5cm,
    top=4cm,
    bottom=4cm
]{geometry}
\usepackage{courier}
\usepackage{listings}
\usepackage{tocloft}
\usepackage{xcolor}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amsfonts}
\usepackage{fancyhdr}
\usepackage{enumitem}

% \usepackage{palatino}

% \renewcommand{\familydefault}{\sfdefault}

\lstset{basicstyle=\footnotesize\ttfamily, breaklines=true, numbers=left, language=Python, tabsize=2, aboveskip=15pt, belowskip=15pt}

% Floor and Ceiling
\newcommand\floor[1]{\lfloor#1\rfloor}
\newcommand\ceil[1]{\lceil#1\rceil}

% Github style syntax highlighting
\newcommand{\highlight}{\colorbox{orange!60}}

% Does anyone like bullet points?
\def\labelitemi{--}

\setlength{\headheight}{15.0pt}
\pagestyle{fancy}
\lhead{6.046 Problem Set 1: DNC}
\chead{Jing Lin}
\rhead{Nov. 24 2017}

\begin{document}
\subsection*{Problem 1-1 Restaurant Location}
\par{\textit{Drunken Donuts}, a new wine-and-donuts restaurant chain, wants to build restaurants on many street corners with the goal of maximizing total profit.}
\par{The street network is described as an undirected graph $G = (V, E)$, where the potential restaurant sites are the vertices in the graph. Each vertex $u$ has a nonnegative integer value $p_u$, which describes the potential \textbf{profit} of site $u$. Two restaurants cannot be built on adjacent vertices (to avoid self-competition). Design an algorithm that outputs the chosen set $U \subseteq V$ of sites that maximize the total profit $\sum_{u \subseteq U}p_u$.}
\begin{enumerate}[label=\textbf{\arabic*}]
	\item Consider the following greedy strategy. Choose the highest profit vertex $u_0$ in the tree (breaking ties according to some order on vertex names) and put it into $U$. Remove $u_0$ from further consideration, along with all of its neighbors in $G$. Repeat until no further vertices remain. Give a counterexample to show that this algorithm does not always give a restaurant placement with the maximum profit. \\
	\par{\textbf{Solution:} Consider the counter example $(1, 1, 1, 2)$ with the $2$ connected to the ones. The strategy would output $2$, when the maximum profit is $3$.}
	\item Give an efficient algorithm to determine a placement with maximum profit. \\
	\par{\textbf{Solution:} Pick any vertex, denote it as $u_0$, traverse the tree with DFS and sort the vertices based on their completion times into an array $N$. Children nodes appear before their parents. For each vertex, create two functions, $A(v)$ and $B(v)$ which represent the maximum profit at the subtree rooted at $v$ including and excluding $v$.}
	\par{For a leaf node v}
	\par{$A(v) = p_v$, and $B(v) = 0$}
	\par{For non-leaf nodes}
	\par{$A(v) = p_v + \sum_{u \in v.children}B(u)$}
	\par{$B(v)$ is $\sum_{u \in v.children}\max{(B(u), A(u))}$}
	\par{For each node in $N$, compute $A(v_i)$ and $B(v_i)$. The final node corresponds to the root of the subtree, at which point the maximum strategy is $\max{(A(v_n), B(v_n))}$.}
	\par{\textbf{Correctness:} Consider the proof via induction.}
	\par{In the base case with one node $v$, we know that the maximum value is $p_v$ which for a leaf node is $\max{(A(v), B(v))} = \max{(p_v, 0)} = p_v$.}
	\par{Inductive hypothesis: consider the situation where we have $n + 1$ nodes. There are two combinations to consider, including and excluding the additional node $u$. The maximum profit must be one of the two combinations. $N$ guarantees that children nodes are processed before their parents, so the $A(v)$ and $B(v)$ are accurate.}
	\par{\textbf{Timing:} The algorithm runs in $O(V)$ time. Trees have $\leq V - 1$ edges. Therefore, traversing $G$ and sorting the nodes by their reverse finishing times takes $O(V)$ time. The time it takes to find $A(v)$ and $B(v)$ is also bounded by the $O(V)$ number of edges from each vertex. Therefore, the entire algorithm runs in $O(V)$.}
\end{enumerate}


\section*{Median Finding Algorithms}
\par{There are a few approaches to finding the median of a list.  One approach is to sort the elements of the list and then compute the position of the median. That approach takes $\Theta(n \log{n})$ time.}
\par{Another approach inspired by quicksort is to find a pivot, and recurse on a subarray based on the rank of the pivot. The corresponding pseudocode is presented below.}

\begin{lstlisting}
	SELECT(S, i):
		Pick x in S cleverly
		Compute k = rank(x)
		B = Set of elements in S, such that elements < x
		C = Set of elements in S, such that elements > x
		if k == i:
			return x
		else if k > i:
			return SELECT(B, i)
		else if k < i:
			return SELECT(C, i - k)	
\end{lstlisting}

\par{The intuition here is if we can eliminate either B or C cleverly, and prove that the subarray we recurse on is always a fraction of the previous array, the algorithm will run in $\Theta(n)$ time.}

Without a clever strategy, if we’re selecting the $n - 1$ element and $k = 1$ every time, then we’re eliminating only one element on every iteration. There would be $n$ iterations, each taking $\Theta(n)$ time to partition the subarray. Therefore, the algorithm would run in $\Theta(n^2)$.

\subsubsection*{Clever Partitioning}
\begin{enumerate}
\item Arrange $S$ into columns of size 5 ($\ceil{n / 5}$ columns).
\item Sort each column in linear time (top down)
\item Find the medians of medians as $x$.
\end{enumerate}

\par{Half of the $\ceil{n / 5}$ groups contribute at least 3 elements $> x$ except for $1$ trailing group which could contain fewer than $5$ elements and $1$ group containing $x$. Therefore, there are at least $3(\ceil{n / 10} - 2)$ elements $> x$ and at least $3(\ceil{n / 10} - 2)$ elements $< x$.}
\par{The recurrence can now be written as}

\begin{align*}
f(n) = 
\begin{cases} 
	\Theta(1), & \mbox{for } n \leq 140 \\
	T(\ceil{n / 5}) + T(\frac{7n}{10} + 6) + \Theta(n), & \mbox{for } n > 140
\end{cases}
\end{align*}

\subsubsection*{Proof}
\par{Let's prove that $T(n) < cn$ via induction. The base case is when $n \leq 140$, in which case, $T(n) < cn$ for some arbitrarily large $c$.}
\par{Let's now consider the case when $n > 140$. First, we can substitute the base case into our equation. Then we can add an additional $c$ term.}
\begin{align*}
T(n) &\leq c\ceil{n / 5} + c(\frac{7n}{10} + 6) + an \\
&\leq \frac{cn}{5} + c + \frac{7nc}{10} + 6c + an \\
&= cn + (-\frac{cn}{10} + 7c + an)
\end{align*}
\par{It's important to now check for when $(-\frac{cn}{10} + 7c + an) = 0$.}
\begin{align*}
(-\frac{cn}{10} + 7c + an) &\leq 0 \\
\frac{cn}{10} &\geq 7c + an \\
c &\geq \frac{70c}{n} + 10a
\end{align*}
\par{The expression is always true for $n \geq 140$ and $c \geq 20a$.}

\subsubsection*{Code}
\begin{lstlisting}
import math
import random

const = 20
sub_size = 5
def SELECT(S, i):
	if len(S) < const:
		return sorted(S)[i] # assuming rank is zero indexed
	num_medians = int(math.ceil(len(S) / sub_size))
	medians = []
	for curr_index in range(0, len(S), sub_size):
		sub = S[curr_index : min(curr_index + sub_size, len(S))]
		mid_index = int(math.floor(len(sub) / 2))
		medians.append(sorted(sub)[mid_index])

	median = SELECT(medians, int(math.floor(len(medians) / 2)))
	B = []
	C = []
	for element in S:
		if element < median:
			B.append(element)
		elif element > median:
			C.append(element)
	if len(B) == i:
		return median
	elif len(B) > i:
		return SELECT(B, i)
	else:
		return SELECT(C, i - len(B) - 1)

arr = range(40)
random.shuffle(arr)
print(arr)
print(SELECT(arr, 15))
\end{lstlisting}

\end{document}






